\thetitle

### 2021-05-17 paper/paper.tex
minor

\begin{abstract}
  % teilw. von IJPR paper uebernommen
  A major task in hierarchical structured manufacturing planning and control is to determine when to
  release orders to the shop floor. One key parameter is the lead time which is the planned time
  that elapses between the release of an order and its completion. Lead times are usually determined
  based on the flow times, i.e.\@ the measured duration orders take to traverse the production
  system. Traditional order release models assume static lead times, although it has been shown that
  they should be set dynamically to be able to react to the dynamic operational characteristics of
  production system. Therefore, we present a reinforcement learning order release model which sets
  lead times dynamically. The applied reinforcement learning agent is designed especially for
  processes with periodic feedback and high variability. This paper aims at providing a
  proof-of-concept for the proposed algorithm. We compare our method on a three-stage flow-shop
  simulation model not only to static order release techniques, but also to the commonly used deep
  Q-Learning agent. The results show that both reinforcement learning agents have room for
  improvement for queuing systems with low variability, that our proposed method usually outperforms
  deep Q-Learning and finally for systems with high utilisation that the proposed reinforcement
  learning algorithm performs very well, outperforming all other approaches.
  %
  % We use a  to compare the algorithm
  % to % and show that our approach outperforms
  % well-known order release mechanisms. We show that in the current version our proposed model using
  % reinforcement learning outperforms some, but not all other tested approaches.
  % EVTL: especially for scenarios with...
  \keywords{operations research, production planning, order release, machine learning, reinforcement
    learning}
\end{abstract}

@@

One method for which the future system state is by design considered a main driver of decision
making % that is designed to incorporate the future system state
is reinforcement learning \citep[e.g.][]{sutton1998introduction}, one of the main three
subcategories of machine learning. % It stems from dynamic programming the future
% system state is by design considered as a main driver of decision making in the current period.
Reinforcement learning is an optimisation technique that stems from dynamic programming, therefore
also seeks to compute the best stationary policy for a given problem. The idea is to evaluate the
current system state according to the direct feedback (reward), as well as the previously learned
state value of the consecutive state. This forms a cyclic constraint problem with as many
constraints and variables as there are system states. The solution is found through iteration and
due to the transitivity property of the state value update.
% as in dynamic programming is that the process and the current state and future states (or
% state-action pairs) of an underlying Markov Decision Process (MDP).
The advantage of reinforcement learning over dynamic programming is that (i) the problem space is
explored by an agent and thus only expectantly interesting parts of the problem space need to be
assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary as the
states are evaluated by consecutively observed states solely.

There are two main types of reinforcement learning algorithms known in literature. The commonly
applied discounted reinforcement learning method
\citep[e.g.][]{sutton1998introduction,mnih2015human,mnih2016asynchronous}, in which the future
system state is discounted by a factor \(0 < \gamma < 1\) to prevent an infinite growth of the state
evaluations, and average reward reinforcement learning
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,
  schwartz1993reinforcement,
  Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
%
% Like discounted reinforcement learning also average reward reinforcement learning is based on an
% oracle function, in our case the accumulated costs of a period, to assess the decisions taken by
% the agent. By repeatedly choosing different actions the agent examines the problem space and rates
% possible actions for any observed state. The advantage of average reward reinforcement learning
The advantage of the later % over the widely applied discounted reinforcement learning framework
is that the underlying optimisation technique is theoretically able to find better policies
% This yields from the fact that in the standard framework each state is assessed independently and
% by a single value. To be more precise, average reward reinforcement learning
by splitting up the evaluation of each state into infinitely many parts. First, the average reward
per step, second, a bias value that specifies expected reward that is additionally accumulated when
traversing from the current state to the optimal path, and third, an infinite number of values of
which each includes the occurrence of the rewards and in which magnitude along the path
\citep{Howard64,Puterman94,
  Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}. % Thus,
% while in average reward reinforcement learning these terms are learned separately
To expectantly accumulate the most reward one has to choose the action that leads to the state with
highest value of the parts, on ties selecting actions following the order above. That is, the
highest evaluation of the average reward, the bias value, the third component, the fourth component,
etc. The formulations for computing these values form nested constraints, so for instance the
current estimate of the average reward is used to compute the bias value. Therefore, a slight
approximation error may cause harmful damage in subsequent constraints, which makes average reward
reinforcement computationally very intensive. Worsening the issue, \citet[p.343ff]{Puterman94} shows
that an additional set of nested constraints have to be solved to gain the correct solution.

In contrast the standard discounted framework estimates only one single scalar value consisting of
the addition of these subterms, where however, the average reward is scaled by \(1/(1-\gamma)\).
%
To collect as much reward as possible, for instance to gain the highest expected profit, the
discount factor \(\gamma\) has to be very close to one, e.g. \(0.99\).
%
% As the discount factor \(\gamma\) is usually set very
% close to \(1\), e.g. \(0.99\), which
Such a parameterisation leads to the fact that the subterm including the average reward dominates
the other two.
%
Further, in commonly applied unichain (and thus ergodic) MDPs the average reward per step is equal
over all states. Thus, independently assessing it is not only computationally unwisely, but also
problematic when they are iteratively approximated. For this reasons most applications of discounted
reinforcement learning in literature evaluate their algorithms on processes that have an average
reward of about zero \citep{schneckenreither2020average}. \citet{schneckenreither2020average} shows
that tabular based discounted reinforcement learning agents are unable to find even basic policies
when confronted with processes that have a small average reward per step. Unfortunately, when
optimising for the highest profit (or the lowest cost) in the area of operations management, the
most intuitive and clear way of reward feedback to the agent is presenting it the profit that is
accumulated every period.

Therefore, we develop, adapt and enhance an algorithm that uses only one scalar value for the
estimation of the average reward over all states. Further it estimates for every state-action pair
the bias value and the sum of all other state value parts, usually referred to as error term. The
later values are adjusted by the average reward.
% This, combined with the iterative evaluation and the independently assessing of the average reward
% for each state lets standard discounted reinforcement learning struggle to find good policies.
% Thus, this incautious combining of different kinds of state values as done in discounted
% reinforcement learning leads to the problems that (i) the average reward, which is equal for all
% states of usually investigated unichain MDPs, is dominating and thus diluting the bias values, and
% (ii) the state values are deteriorated by the error term that is only imposed due to the
% discounting technique.
%
%
Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an
average reward adjusted reinforcement learning algorithm to dynamically set lead time for order
release. In contrast to the aforementioned works on average reward reinforcement learning our
algorithm incorporates the optimisation of not only the average reward over time, but also the bias
values, which is an important adaption in highly stochastic systems. Only the work by
\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
integrates this second-level refinement optimisation. However, their average reward reinforcement
learning algorithm requires the selection of a reference state to prevent an infinite increase of
state values. This results from the lack of feedback in the iterative process of optimising
different sets of nested constraints. Furthermore, they asses the average reward independently for
each state.

@@

Finally, we provide a proof-of-concept experimental section by testing the proposed algorithm on a
three stage flow shop with six product types, which provides ample evidence of the viability of the
approach.
% and are compared to static order release
% mechanisms. Furthermore, we show that the advantage of average reward reinforcement learning for the
% problem structures encountered in the context of production and logistics management by comparing
% the results to the discounted reinforcement learning
% framework.
The highly optimised value-iteration algorithm with improved n-step learning, small action-based
replay memories and concurrent workers infers order release policies which, especially in case of
high utilisation, outperform not only the static order release mechanisms, but also the commonly
applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.
% With regard to practical implications we are confident that decision support
% tools based on average reward reinforcement learning increase the decision quality of human
% planners.
%
In summary our contributions are as follows:
\begin{itemize}
\item We propose a novel \textit{average reward adjusted deep reinforcement learning algorithm} \ARA{}.
  This is the first average reward adjusted reward learning algorithm that operates using a
  artificial neural network as function approximation. It includes several known extensions like
  n-step learning \citep[e.g.][]{mnih2016asynchronous} or experience replay memory
  \citep[e.g.][]{mnih2015human}, which we advanced to prevent catastrophic forgetting by
  action-based selection of experiences.
\item We establish a action reduction method, named \textit{independent agents}, as a subset of the
  action space is treated as if there were multiple independent agents, although they share the same
  state. It reduces the growth of the number of possible actions from a exponential growth to a
  linear one. This is especially useful for product type based lead time management, as each agent
  is responsible for the lead time of one product type.
\item Using a three-stage flow shop model we show in an \textit{experimental evaluation} that the
  proposed algorithm outperforms standard Q-Learning and well established order release methods,
  especially for systems with high variability.
\end{itemize}

@@

This section is split into several parts. First we discuss lead time management literature, then
reinforcement learning applications to lead time management, third we briefly review the relevant
notation to reinforcement learning and finally we present the relevant literature of reinforcement
learning in the rest of the section. The later includes a discussion of discounted reinforcement
learning, average reward reinforcement learning, the Laurent series expansion of the state values,
and finally average reward adjusted reinforcement learning.

@@

Setting lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
release mechanisms use static or fixed lead times to address the order release problem.
%
However, lead times should be set dynamically due to the non-linear relationship between utilisation
and flow times \citealt{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
system performance. For instance, a solely reactive approach can result in the lead time syndrome.
%
The \textit{lead time syndrome} describes a vicious cycle where increasing flow times perpetually
inflate the lead times which leads to worse performance (see e.g., \citealt{Mather1978,
  knollmann2013control, Selcuk2006}).
%
%
\citet{schneckenreither2020order} have established following categorisation of dynamic lead time
management approaches:

\begin{itemize}
\item \textsf{Reactive lead time management} sets lead times by reacting on earlier flow times
  (e.g., \citealt{enns2004work, Selcuk2006}), i.e.\@ data from the past.
\item \textsf{Proactive lead time management} may incorporate \textit{past} data and the
  \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983, ChungHuang2002}).
\item \textsf{Predictive lead time management} may incorporate \textit{past} data and the
  \textit{current} system state, and uses an anticipated \textit{future} system state to prevent
  arising issues of future periods (e.g.
  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}).
\end{itemize}

@@

\paragraph{Reactive LTM.}
%
%% Reactive Due Date Ass
Early due date assignment approaches \citep[e.g.][]{conway1967} set the due dates according to the
processing time (Total Work Content rule, TWK rule) or the number of operations.
%
%% Reactive Order Release
For order release \citet{enns2004work} investigate the use of a reactive approach with exponential
smoothing of flow times to set lead times for two-stage distribution requirements planning. Although
it can be considered an exploratory research paper the main conclusions are that dynamically setting
planned lead times using exponential smoothing works well, but that the reactions are delayed. The
later is due to the nature of the flow time measurement only being possible for finished orders.
%
Another reactive approach uses exponential smoothing to forecast flow times to set lead times for a
capacitated multi-stage make-to-stock system \citep{Selcuk2006}. Their approach performs better than
fixed lead time techniques for most experiments. However, they also exhibit the LTS\@.

\paragraph{Proactive LTM.}
%
%% Proactive Due Date Assignment
The TWK due date assignment rules have been extended by several researchers to proactive methods
\citep{eilon1976, kanet1986, Enns1995}. As flow times depend on the workload and machine capacity in
the shop workload dependent \citep{Bertrand1983} due date assignment rules have been developed. The
basic idea is to estimate the flow time of an order on its arrival by measuring the number of orders
queued on the corresponding routing, which can be analytically \citep[e.g.][]{ChungHuang2002}, by
the use of a simulation model \citep[e.g.][]{vig1991}, statistical methods \citep[e.g.][]{Tai2012}
or using hybrid approaches \citep[e.g.][]{Kaplan1993}. Furthermore several proactive due date
assignment methods that incorporate neural networks exist \citet{Philipoom1994, Philipoom1997,
  HsuSha2004,Patil2008,ChangetalFuzzyDDset2008}.
%
%% Proactive Order Release
In the case of order release the analytical papers by \citet{Selcuk2009} and \citet{Selcuk2013} use
the WIP level to quote lead times. In the former they show that the less frequently updating the
lead time decreases the nervousness and improves the performance, while in the latter that
dynamically updating lead times performs much better than static lead times.
%
% MIL (Modified infinite loading)
MIL extends the static BIL procedure to integrate shopfloor congestion by subtracting number of
queued jobs on the corresponding routing from the due date
\citep{RagatzMabert1988,wisner1995review}. For non-due-date oriented dispatching rules, their
approach has a positive effect on total cost performance, while for due-date oriented dispatching
this holds only if the due dates are loose. For all other cases the release algorithm has no
negative effect on performance. Variants of this rule-based method exist
\citep[e.g.][]{park1989job,kim1995evaluating}.
% Bertrand 2002
% \citet{bertrand2002investigation} investigate different ways of releasing orders, assigning lead
% times and operation due dates to assembly orders. In their work a due date is set once the order
% arrives and then a release time calculated by subtracting the lead time from it. In one approach the
% lead time calculation is the sum of the operation times plus a factor consisting of the waiting time
% allowance per operation times the number of operations. They use the average operation waiting time
% as waiting time allowance factor. In the other approaches the orders are released using the
% \ac{imre} policy. In terms of order release they find that all work orders of an assembly order
% should be released at the same time.
%
A machine learning approach is presented by \citet{Lee1997}. They combine a genetic algorithm with a
decision tree to generate sequencing rules in the order pool for release. In the tested job shops
the CONWIP rule determines the release of the orders and a decision tree to find the order release
sequence. Their method leads to significant improvements over conventional single release and
dispatching approaches.

@@

For scheduling \citet{wang2007reinforcement} present a predictive lead time management approach
using reinforcement learning. Their algorithm uses multiple agents to forecast the processing time
of each job for routing decisions. They show on a 10 machine job shop that the RL algorithm performs
equally or better than tested heuristics.
%
In case of order release, \citet{PaterninaArboleda2001} introduce a predictive order release model
using reinforcement learning in a single product, four-station serial flow line. They compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP). The
algorithm decides on whether or not to release an order after each system change, the completion of
an operation of any order at any stage or a new order arrival, and assumes that any unsatisfied
demand is lost. Thus, they use a continuous order release method although in practice order release
decisions often need to be made on a periodical basis, e.g.\@ daily (see
\citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their
tabular based reinforcement learning agent.
%
Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the syndrome. Their model outperforms static and reactive lead time
management approaches, especially for scenarios with high utilisation.

@@

Over the past decades reinforcement learning has been applied to various problems, for which
astonishing results have been reported. E.g.\@ only recently \citet{mnih2015human} have presented a
novel value-iteration reinforcement learning agent which exceeds human-level abilities in playing
many classic Atari games~\citep{bellemare2012investigating}. Further, \citet{mnih2016asynchronous}
present improved results with asynchronous actor-critic reinforcement learning. Also games like
Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have been mastered with
superhuman performance by \textit{tabula rasa} reinforcement learning agents. Furthermore, the
method has also been applied in the setting of manufacturing system, e.g.\@ to improve the ramp-up
process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate dispatching rules
\citep{zhang1995reinforcement,wang2005application} or scheduling \citep{zhang1995reinforcement,
  waschneck2018optimization}.
%
However, all these applications use discounted reinforcement learning and are either designed to
investigate a rather simple MDP, e.g.\@ by selecting heuristics instead of optimising the underlying
problem itself, or by mapping the actual objective in a reward function that is approximately \(0\)
on average over time.
%
This is due to the fact that state value is largely composed of a term defined by the policy's
average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the state
values and thus decrease the solution quality, as can for instance be observed in
\citet{SchneckenreitherHaeussler2019} and \citet{gijsbrechts2018can}.
%

>>>>

\begin{abstract}
  % teilw. von IJPR paper uebernommen
  A major task in hierarchically structured manufacturing planning and control is to determine when
  to release orders to the shop floor. One key parameter is the lead time which is the planned time
  that elapses between the release of an order and its completion. Lead times are usually determined
  based on the flow times, i.e.\@ the measured duration orders take to traverse the production
  system. Traditional order release models assume static lead times, although it has been shown that
  they should be set dynamically to be able to react to the dynamic operational characteristics of
  production system. Therefore, we present a reinforcement learning order release model which sets
  lead times dynamically. The applied reinforcement learning agent is designed especially for
  processes with periodic feedback and high variability. This paper aims at providing a
  proof-of-concept for the proposed algorithm. We compare our method on a three-stage flow-shop
  simulation model not only to static order release techniques, but also to the commonly used deep
  Q-Learning agent. The results show that both reinforcement learning agents have room for
  improvement for queuing systems with low variability, that our proposed method usually outperforms
  deep Q-Learning and finally for systems with high utilisation that the proposed reinforcement
  learning algorithm performs very well, outperforming all other approaches.
  %
  % We use a  to compare the algorithm
  % to % and show that our approach outperforms
  % well-known order release mechanisms. We show that in the current version our proposed model using
  % reinforcement learning outperforms some, but not all other tested approaches.
  % EVTL: especially for scenarios with...
  \keywords{operations research, production planning, order release, machine learning, reinforcement
    learning}
\end{abstract}

@@

One method for which the future system state is by design considered a main driver of decision
making % that is designed to incorporate the future system state
is reinforcement learning \citep[e.g.][]{sutton1998introduction}, one of the main three
subcategories of machine learning. % It stems from dynamic programming the future
% system state is by design considered as a main driver of decision making in the current period.
Reinforcement learning (RL) is an optimisation technique that stems from dynamic programming,
therefore also seeks to compute the best stationary policy for a given problem. The idea is to
evaluate the current system state according to the direct feedback (reward), as well as the
previously learned state value of the consecutive state. This forms a cyclic constraint problem with
as many constraints and variables as there are system states. The solution is found through
iteration and due to the transitivity property of the state value update.
% as in dynamic programming is that the process and the current state and future states (or
% state-action pairs) of an underlying Markov Decision Process (MDP).
The advantage of RL over dynamic programming is that (i) the problem space is
explored by an agent and thus only expectantly interesting parts of the problem space need to be
assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary as the
states are evaluated by consecutively observed states solely.

There exist two main types of deep RL algorithms in literature. The commonly
applied discounted RL method
\citep[e.g.][]{sutton1998introduction,mnih2015human,mnih2016asynchronous}, in which the future
system state is discounted by a factor \(0 < \gamma < 1\) to prevent an infinite growth of the state
evaluations, and average reward RL
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,
  schwartz1993reinforcement,
  Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
%
% Like discounted reinforcement learning also average reward reinforcement learning is based on an
% oracle function, in our case the accumulated costs of a period, to assess the decisions taken by
% the agent. By repeatedly choosing different actions the agent examines the problem space and rates
% possible actions for any observed state. The advantage of average reward reinforcement learning
The advantage of the later % over the widely applied discounted reinforcement learning framework
is that the underlying optimisation technique is theoretically able to find better policies
% This yields from the fact that in the standard framework each state is assessed independently and
% by a single value. To be more precise, average reward reinforcement learning
by splitting up the evaluation of each state into infinitely many components. First, the average
reward per step, second, a bias value that specifies expected reward that is additionally
accumulated when traversing from the current state to the optimal path, and third, an infinite
number of values of which each includes the occurrence of the rewards and in which magnitude along
the path \citep{Howard64,Puterman94,
  Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}. % Thus,
% while in average reward reinforcement learning these terms are learned separately
To expectantly accumulate the most reward one has to choose the action that leads to the state with
highest value of the components following the order above, on ties
selecting % within the best actions
according to the next component. That is, the highest evaluation of the average reward, the bias
value, the third component, the fourth component, etc. The formulations for computing these values
form nested constraints, so for instance the current estimate of the average reward is used to
compute the bias value. Therefore, a slight approximation error may cause harmful damage in
subsequent constraints, which makes average reward reinforcement computationally very intensive.
Worsening the issue, \citet[p.343ff]{Puterman94} shows that the constraint problem has to be
specified to solve one more component as originally intended to gain correct solution.

In contrast the standard discounted framework estimates only one single scalar value consisting of
the addition of these components, where however, the average reward is scaled by \(1/(1-\gamma)\).
%
But in order to collect as much reward as possible the discount factor \(\gamma\) has to be very
close to one, for instance \(0.99\).
%
% As the discount factor \(\gamma\) is usually set very
% close to \(1\), e.g. \(0.99\), which
Such a parameterisation leads to the fact that the subterm including the average reward dominates
the other two.
%
Further, in commonly applied unichain (and thus ergodic) MDPs the average reward per step is equal
over all states. Thus, independently assessing it is not only computationally unwisely, but also
problematic as they are iteratively approximated. For this reasons most applications of discounted
RL in literature evaluate their algorithms on processes that have an average
reward of about zero \citep{schneckenreither2020average}. \citet{schneckenreither2020average} shows
that tabular based discounted RL agents are unable to find even basic policies
when confronted with processes that have a small average reward per step. Unfortunately, when
optimising for the highest profit (or the lowest cost) in the area of operations management, the
most intuitive and clear way of reward feedback to the agent is presenting it the profit that is
accumulated every period.

Therefore, we develop an algorithm that uses only one scalar value for the estimation of the average
reward over all states. Further it estimates for every state-action pair the bias value and the sum
of all other state value components, usually referred to as error term. The later values are
adjusted by the average reward.
% This, combined with the iterative evaluation and the independently assessing of the average reward
% for each state lets standard discounted reinforcement learning struggle to find good policies.
% Thus, this incautious combining of different kinds of state values as done in discounted
% reinforcement learning leads to the problems that (i) the average reward, which is equal for all
% states of usually investigated unichain MDPs, is dominating and thus diluting the bias values, and
% (ii) the state values are deteriorated by the error term that is only imposed due to the
% discounting technique.
%
%
Thus, as opposed to the commonly applied discounted RL algorithm we use an
average reward adjusted RL algorithm to dynamically set lead time for order
release. In contrast to the aforementioned works on average reward RL our
algorithm incorporates the optimisation of not only the average reward over time, but also the bias
values, which is an important adaption in highly stochastic systems. Only the work by
\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
integrates this second-level refinement optimisation. However, their average reward reinforcement
learning algorithm requires the selection of a reference state to prevent an infinite increase of
state values. This results from the lack of feedback in the iterative process of optimising
different sets of nested constraints. Furthermore, they asses the average reward independently for
each state.

@@

Finally, we provide a proof-of-concept experimental section by testing the proposed algorithm on a
three stage flow shop with six product types, which provides ample evidence of the viability of the
approach.
% and are compared to static order release
% mechanisms. Furthermore, we show that the advantage of average reward reinforcement learning for the
% problem structures encountered in the context of production and logistics management by comparing
% the results to the discounted reinforcement learning
% framework.
The highly optimised value-iteration algorithm with improved n-step learning, small action-based
replay memories and concurrent workers infers order release policies which, especially in case of
high utilisation, outperform not only the static order release mechanisms, but also the commonly
applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.
% With regard to practical implications we are confident that decision support
% tools based on average reward reinforcement learning increase the decision quality of human
% planners.
%
In summary our contributions are as follows:
\begin{itemize}
\item We propose a novel \textit{average reward adjusted deep reinforcement learning algorithm}
  \ARA{}. This is the first average reward adjusted reward learning algorithm that operates using a
  artificial neural network as function approximation. It includes several known extensions like
  n-step learning \citep[e.g.][]{mnih2016asynchronous} or experience replay memory
  \citep[e.g.][]{mnih2015human}, which we aditionally advanced to prevent catastrophic forgetting by
  action-based selection of experiences.
\item We establish a action reduction method, named \textit{independent agents}, as a subset of the
  action space is treated as if there were multiple independent agents, although they share the same
  state. It reduces the growth of the number of possible actions from a exponential growth to a
  linear one. This is especially useful for product type based lead time management, as each agent
  is responsible for the lead time of one product type.
\item Using a three-stage flow shop model we show in an \textit{experimental evaluation} that the
  proposed algorithm outperforms standard Q-Learning and well established order release methods,
  especially for systems with high variability.
\end{itemize}

@@

This section is split into several parts. First we discuss lead time management literature, then
RL applications to lead time management, third we briefly review the relevant
notation to RL and finally we present the relevant literature of reinforcement
learning in the rest of the section. The later includes a discussion of discounted reinforcement
learning, average reward RL, the Laurent series expansion of the state values,
and finally average reward adjusted RL.

@@

Setting lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
release mechanisms use static or fixed lead times to address the order release problem.
%
However, lead times should be set dynamically due to the non-linear relationship between utilisation
and flow times \citep{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
system performance. For instance, a solely reactive approach can result in the lead time syndrome.
%
The \textit{lead time syndrome} describes a vicious cycle where increasing flow times perpetually
inflate the lead times which leads to worse performance (see e.g., \citealt{Mather1978,
  knollmann2013control, Selcuk2006}).
%
%
\citet{schneckenreither2020order} have established following categorisation of dynamic lead time
management approaches:

\begin{itemize}
\item \textsf{Reactive lead time management} sets lead times by reacting on earlier flow times
  (e.g., \citealt{enns2004work, Selcuk2006}), i.e.\@ data from the \textit{past}.
\item \textsf{Proactive lead time management} may incorporate \textit{past} data and the
  \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983, ChungHuang2002}).
\item \textsf{Predictive lead time management} may incorporate \textit{past} data and the
  \textit{current} system state, and uses an anticipated \textit{future} system state to prevent
  arising issues of future periods (e.g.
  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}).
\end{itemize}

@@

\paragraph{Reactive LTM.}
%
%% Reactive Due Date Ass
Early due date assignment approaches \citep[e.g.][]{conway1967} set the due dates according to the
processing time (Total Work Content rule, TWK rule) or the number of operations.
%
%% Reactive Order Release
For order release \citet{enns2004work} investigate the use of a reactive approach with exponential
smoothing of flow times to set lead times for two-stage distribution requirements planning. Although
it is considered an exploratory research paper the main conclusions are that dynamically setting
planned lead times using exponential smoothing works well, but that the reactions are delayed. The
later is due to the nature of the flow time measurement only being possible for finished orders.
%
Another reactive approach uses exponential smoothing to forecast flow times to set lead times for a
capacitated multi-stage make-to-stock system \citep{Selcuk2006}. Their approach performs better than
fixed lead time techniques for most experiments. However, they also exhibit the LTS\@.

\paragraph{Proactive LTM.}
%
%% Proactive Due Date Assignment
The TWK due date assignment rules have been extended by several researchers to proactive methods
\citep{eilon1976, kanet1986, Enns1995}. As flow times depend on the workload and machine capacity in
the shop \citep{Bertrand1983} workload dependent due date assignment rules have been developed. The
basic idea is to estimate the flow time of an order on its arrival by measuring the number of orders
queued on the corresponding routing, which can be analytically \citep[e.g.][]{ChungHuang2002}, by
the use of a simulation model \citep[e.g.][]{vig1991}, statistical methods \citep[e.g.][]{Tai2012}
or using hybrid approaches \citep[e.g.][]{Kaplan1993}. Furthermore several proactive due date
assignment methods that incorporate neural networks exist \citep{Philipoom1994, Philipoom1997,
  HsuSha2004,Patil2008,ChangetalFuzzyDDset2008}.
%
%% Proactive Order Release
In the case of order release the analytical papers by \citet{Selcuk2009} and \citet{Selcuk2013} use
the WIP level to quote lead times. In the former they show that the less frequently updating the
lead time decreases the nervousness and improves the performance, while in the latter that
dynamically updating lead times performs much better than static lead times.
%
% MIL (Modified infinite loading)
Modified infinite loading extends the static Backward Infinte Loading (BIL) procedure to integrate
shopfloor congestion by subtracting number of queued jobs on the corresponding routing from the due
date \citep{RagatzMabert1988,wisner1995review}. For non-due-date oriented dispatching rules, their
approach has a positive effect on total cost performance, while for due-date oriented dispatching
this holds only if the due dates are loose. For all other cases the release algorithm has no
negative effect on performance. Variants of this rule-based method exist
\citep[e.g.][]{park1989job,kim1995evaluating}.
% Bertrand 2002
% \citet{bertrand2002investigation} investigate different ways of releasing orders, assigning lead
% times and operation due dates to assembly orders. In their work a due date is set once the order
% arrives and then a release time calculated by subtracting the lead time from it. In one approach the
% lead time calculation is the sum of the operation times plus a factor consisting of the waiting time
% allowance per operation times the number of operations. They use the average operation waiting time
% as waiting time allowance factor. In the other approaches the orders are released using the
% \ac{imre} policy. In terms of order release they find that all work orders of an assembly order
% should be released at the same time.
%
A machine learning approach is presented by \citet{Lee1997}. They combine a genetic algorithm with a
decision tree to generate sequencing rules in the order pool for release. In the tested job shops
the CONWIP rule determines the release of the orders and a decision tree to find the order release
sequence. Their method leads to significant improvements over conventional single release and
dispatching approaches.

@@

For scheduling \citet{wang2007reinforcement} present a predictive lead time management approach
using RL. Their algorithm uses multiple agents to forecast the processing time
of each job for routing decisions. They show on a 10 machine job shop that the RL algorithm performs
equally or better than tested heuristics.
%
In case of order release, \citet{PaterninaArboleda2001} introduce a predictive order release model
using RL in a single product, four-station serial flow line. They compare its
performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP). The
algorithm decides on whether or not to release an order after each system change, the completion of
an operation of any order at any stage or a new order arrival, and assumes that any unsatisfied
demand is lost. Thus, they use a continuous order release method although in practice order release
decisions often need to be made on a periodical basis, e.g.\@ daily (see
\citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their
tabular based RL agent.
%
Then, \citet{Schneckenreither2019} use several different RL algorithms to make
periodic order release decisions for a flow shop production system. The algorithm directly sets lead
times for each product type, which are then used to release the orders in the order pool. They show
that their approach outperforms static order release mechanisms by yielding lower costs, lateness
and standard deviation of lateness, but conclude that research using average reward reinforcement
learning methods harbor optimisation potential for the order release problem.
%
And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
times dynamically using an artificial neural network, which is used to forecast flow times. By
implementing a rolling horizon order release simulation of the proceeding periods they lift their
approach to a predictive lead time management approach, which is able to detect backorders of future
periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
prevent the negative effects of the syndrome. Their model outperforms static and reactive lead time
management approaches, especially for scenarios with high utilisation.

@@

Over the past decades RL has been applied to various problems, for which
astonishing results have been reported. E.g.\@ only recently \citet{mnih2015human} have presented a
novel value-iteration RL agent which exceeds human-level abilities in playing
many classic Atari games~\citep{bellemare2012investigating}. Further, \citet{mnih2016asynchronous}
present improved results with asynchronous actor-critic RL. Also games like
Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have been mastered with
superhuman performance by \textit{tabula rasa} RL agents. Furthermore, the
method has also been applied in the setting of manufacturing system, e.g.\@ to improve the ramp-up
process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate dispatching rules
\citep{zhang1995reinforcement,wang2005application} or scheduling \citep{zhang1995reinforcement,
  waschneck2018optimization}.
%
However, all these applications use discounted RL and are either designed to investigate a rather
simple MDP, e.g.\@ by selecting heuristics instead of optimising the underlying problem itself, or
by mapping the actual objective in a reward function that is approximately zero on average over
time.
%
This is due to the fact that state value is largely composed of a term defined by the current
policys' average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the
state values and thus decrease the solution quality, as can for instance be observed in
\citet{SchneckenreitherHaeussler2019} and \citet{gijsbrechts2018can}.
%

Therefore, most applications that incorporate and directly reflect costs or profit in the reward
function use average reward RL. \citet{aydin2000dynamic} use it in the setting
of scheduling, while in a series of papers Mahadevan et al.\@ investigated several problem domains
starting with simple MDPs
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
After these foundational works for average reward RL they introduced a
continuous time average reward RL algorithm named SMART
\citep{Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
Applications of SMART reach from the optimisation of queuing systems
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies,Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning},
maintenance of an inventory system
\citep{Das99_SolvingSemiMarkovDecisionProblemsUsingAverageRewardReinforcementLearning} to optimising
transfer line in terms of maximizing demand, while keeping inventory levels of unfinished product as
low as possible \citep{Mahadevan98_OptimizingProductionManufacturingUsingReinforcementLearning}.
However, in practise usually decision have to be made on a daily basis
\citep{enns2004work,GeldersvanW1982}. Therefore, we refrain from this adaption and concentrate on
standard MDPs only. Furthermore, often continuous-time semi-MDP problems can be converted through
uniformisation into equivalent discrete time instances
\citep[see][]{Puterman94,bertsekas1995dynamic}.
